Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme\\ \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

### 2020-04-27 _region_.tex
Cambios en apéndice y modificaciones a comentarios

\message{ !name(LME-MPM.tex)}
\documentclass[preprint,12pt,a4paper]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}

@@

\newcommand{\vec}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

@@

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}

@@

@@

\title{Enhanced Material Point Method to face
  dynamic problems: Local-maximum entropy approximation and explicit
  predictor-corrector scheme}

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard \acrshort{MPM} time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an explicit predictor-corrector time
  integration scheme has been proposed. Its superior performance
  mitigates the presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in \acrshort{MPM} 
  occurs when to material points cross computational grid
  boundaries and motivated due to the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \acrfull{lme} approximation schemes a
  robust substitute for the wide range of shape function in
  \acrshort{MPM}. Local \textit{max-ent} approximation may be regarded
  as a \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

\begin{keyword}
  \acrshort{lme} \sep \acrshort{mpm} \sep Explicit predictor-corrector \sep Dynamic problems}
\end{keyword}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}. It popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is  a voxel based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive. In recent years the employ of
spline-lines has gain popularity with the introduction of the B-Spline
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employ of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to \acrshort{mpm} framework with the Conservative Taylor
Least Squares (CTLS) reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}, unfortunately when particles are spread in
a challenging way, the quality of the CTLS approximation decrease
locally.

This document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) as a robust substitute for the wide range of shape
  function in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite element method (FEM) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions~\cite{Arroyo2006}. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)~\cite{Navas2018a} and Li {\it et al.} (2012)~\cite{Li2012} for
Optimal Transportation Meshfree (OTM) method. And more
recently under \acrshort{mpm} framework by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020} but without exploring the benefits of
the regularization parameter \gls{LME-beta}.

This techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using
techniques~\cite{Tran2019e}. These numerical inaccuracies also known
as wiggles are due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises is to add a
nonphysical damping source to the equilibrium equations, this
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing the accuracy of the time integration
scheme. One of the most popular is the \acrfull{igimp}
Charlton {\it et al.} (2017)~\cite{Charlton2017}, more recently Tran \&
Solowski (2019)~\cite{Tran2019e} proposed a generalised-$\alpha$ scheme
for \acrshort{mpm} with promising results but at the expense of increasing
the computational effort. In this paper a less time consuming and high
efficient explicit predictor corrector integration method has been
proposed. It consists in an accommodation of the Newmark
predictor-corrector or central difference explicit (CD). We have
choose this method among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because it simplicity and it good
performance dealing with solids dynamics problems under a mesh-free
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employ of a
suitable combination of the maximum-entropy (or local
\textit{max-ent}) family shape functions, and the proposal of
a explicit predictor-corrector scheme. We illustrate the advantages of
this approach thorough to simple but challenging test cases, like the
propagation of shock waves in a elastic bar or the response of a block of soil
gradually loaded with gravitational forces.

The article is organized as follows. Section \ref{sec:Notation}
briefly reviews the notation here employed. Next, Section \ref{sec:derivation-mpm}
is devoted to present briefly the governing equations of the elastic
problem, the variational formulation and the Galerkin procedure. In Section \ref{sec:epc-algor-mpm} an explicit predictor-corrector
time integration scheme for \acrshort{mpm} is proposed. Section
\ref{sec:local-max-ent} briefly reviews the local \textit{max-ent}
basis functions here employed. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems} applications
to prove the numerical accuracy of the proposed approach are presented. Finally,
conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

\section{The meshfreesubsection procedure}
\label{sec:derivation-mpm}

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. In consequence it
is structured as follows: the governing equations will be introduced
in \ref{sec:The boundary-value-problem}, later the variational statement of
the problem will be presented \ref{sec:variational-formulation}, and
finally a discretization procedure thorough Galerkin is performed in
\ref{sec:Galerkin-procedure}. The method has three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation thorough a FEM-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carries on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstain subscripts
convention is adopted therefore repeated index means addition.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) the \textit{compatibility
  equation} that extracts from \gls{u} the strain field \gls{strain} is,
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
$\varGamma_d$. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. The corresponding
conjugate variable for the strain field if the stress field
\gls{stress}, which satisfies (ii) the \textit{conservation of
  momentum equation}
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type
$\varGamma_n$. Next (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

To write the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for $\vect{u}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative} represents the weak form
formulation of the problem.

In order to arrive to a finite set of equations, in contrast with the
FEM, in \acrshort{mpm} a double discretization procedure is performed as we will
describe here below. First, the continuum domain
$\Omega$ is discretized with a finite sum of material points (in the
following particles), each one
represent a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point $\vec{x}_p$ is defined at the centroid
of each $\Omega_p$, figure~\ref{fig:MPM-discretization}.
Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$, but also the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where we recover the Riemann
integral definition as an addition of a finite set of points, and
their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =

>>>>

\message{ !name(LME-MPM.tex)}\documentclass[preprint,12pt,a4paper]{elsarticle}

\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

\newcommand{\vect}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

@@

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}

@@

\newcommand{\PNA}[1]{
  \textcolor{red}{{#1}}
}

\newcommand{\MMP}[1]{
  \textcolor{blue}{{#1}}
}

\newcommand{\MMS}[1]{
  \textcolor{gree}{{#1}}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{Enhanced Material Point Method to face
  dynamic problems: Local-maximum entropy approximation and explicit
  predictor-corrector scheme\\
  \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}}

@@

\begin{abstract}
  \acrfull{mpm} has arisen in the recent years as an
  alternative to \acrfull{fem} under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard \acrshort{mpm} time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an Explicit Predictor-Corrector time
  integration scheme has been proposed. Its \textcolor{red}{powerful}
  performance mitigates the presence of spurious oscillations with
  minimal dissipation in high frequency problems. Other source of
  numerical noise in \acrshort{mpm} occurs when the material points
  cross computational grid boundaries, being this issue motivated by
  the lack of smoothness of the interpolation functions. This noise
  results in spurious local variations at the material points, where
  strain-stress fields are computed. This could lead to inaccurate
  solutions as well as aborted simulations in the worst cases. To
  overcome it, this document adopts the \acrfull{lme} approximation
  schemes as a robust substitute of the traditional shape function in
  \acrshort{mpm}. \acrshort{lme} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter
  they are able to behave finite element like or as a meshfree
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
\end{abstract}

\begin{keyword}
  \acrshort{lme} \sep \acrshort{mpm} \sep Explicit predictor-corrector \sep Dynamic problems
\end{keyword}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.\PNA{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

\PNA{On the one hand, the main source of instability occurs} when
material points cross cell boundaries. This provoked the development
of other interpolation techniques to overcome this limitation such as
the \acrfull{gimp} method Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrated to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is a voxel-based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies as noticed Steffen {\it et al.}
(2008)\cite{Steffen2008}. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the \acrfull{ddmp} method proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}. Therefore \acrshort{ddmp} a large number of particles is needed for convergence, what
makes the method very expensive. In recent years the employment of
spline-lines as shape functions has gain popularity with the introduction of the B-Spline 
\acrshort{mpm} proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employment of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to the \acrshort{mpm} framework with the \acrfull{ctls}
reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}. Unfortunately, when particles are spread in
a challenging way, the quality of the \acrshort{ctls} approximation decreases locally.

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute\PNA{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas {\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and coworkers~\cite{Wriggers18} with implicit schemes. And more
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}.\PNA{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several challenging test cases\PNA{on the elastic regime}: the
propagation of shock waves in a elastic bar and the response of a block of soil gradually loaded with gravitational forces.

The article is organised as follows. Section \ref{sec:meshfree-methodology}
is devoted to describe the meshfree methodology adopted in this
research, first \acrshort{mpm} procedure is introduced in
\ref{sec:derivation-mpm}, second the explicit predictor-corrector
time integration scheme is presented in \ref{sec:epc-algor-mpm}, and
third \acrshort{lme} approximation scheme will be introduced in
\ref{sec:local-max-ent}. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems} applications
to prove the numerical accuracy of the proposed approach are
presented. Finally, conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

\section{The meshfree methodology}
\label{sec:meshfree-methodology}

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method is composed by three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition.\PNA{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied.\PNA{First (i) the \textit{compatibility
  equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type \gls{dirichlet-boundary}. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected.\PNA{Therefore, the stress field \gls{stress} will be considered the corresponding conjugate variable for the strain field, being the one} which satisfies (ii) the \textit{conservation of
  momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor:
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\PNA{In this research, plane strain Linear Elasticity has been considered. Thus the constitutive tensor, \tens{D}, is the well known linear elastic one.} The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \};
\end{equation}
satisfying that the Cauchy sequences are convergent in \gls{domain} as well:
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for $\vect{u}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}\\
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
formulation of the problem.
